Community-Centric Innovation

### 2014-03-01 paper.tex
proofreading

Probably the best example of user-centric networking are community networks.
Community networks are deployed by the same users that use them.
Consequently, they are carefully tailored to satisfy the users' needs.
This needs are not necessarily only about connectivity.
The users might be interested in learning about the network, and in researching  new services, applications and protocols.

@@

Another aspect that can draw the attention to community networks is privacy.
As commercial networks are involved in mass surveillance programs, the citizens see in community networks a way to opt out from privacy-invading campaigns.
Finally, community networks are also a means of getting around censorship.
Totalitarian regimes build censorship into the networks to control which information is distributed among the population.
Filtering the information and Internet blackouts have been used as ways for hindering social movements and placate population unrest.

@@

A core principle of community networks is precisely the sharing of the network as a commons good.
This is particularly relevant in the case of wireless networks that use the ISM (industrial, scientific, medical) bands.
This bands contain a small fraction of the spectrum that is available without requiring a licence. 
In practice, these bands are a commons good that should be carefully managed by the community to use it effectively without creating unnecessary interference.
Good practices include the use of directive antennas and the limitation of the radiated power.

@@

There is a large number of community networks and they are very heterogeneous \cite{avonts2013qbe}.
They use different technologies such as wireless point-to-point links, wireless multi-point to multi-point links, ADSL or fiber.
Even in the case in which two community networks use the same layer-2 technology, it is still likely that they use different hardware, firmware or protocols.
The sizes of community networks also vary greatly and can range from few tenths of nodes to tenths of thousands.
The amount of documentation, supporting applications and organizational structure is also very different in different community networks.

Despite the disparity of models for community networks, there is an aspect that is common to all of them.
All these network are user-centric.
They are planned, financed, deployed and maintained by the same users that benefit from them.

@@

Peer-to-peer networks are extremely resilient.
There is no single point of failure or central entity that can be attacked to take down the network.
As there are thousands of nodes spread across the network, the currency can survive as long as the underlying data network survives.
The risks of blockades or political interference are greatly mitigated.
Many Internet organizations such as Wikileaks and Wordpress accept donations and payments in cryptocurrency.
In some cases the reason to bypass government or paying networks blockades and in others the motivation is to operate in countries with lacking payment networks.

It is in areas lacking effective financial and banking systems where cryptocurrencies have a potential for greater impact.
Just like some areas leapfrogged from no telecommunication services to wireless handheld Internet enabled devices, it is possible that these same areas that currently use mobile calls minutes as a currency transition directly to the latest currency technology which is cryptocurrency.
Many countries suffer from chronic hyperinflation problems that could be solved by moving to a cryptocurrency in which the inflation rate (if any) can be chosen by the community and embedded in the protocol.

@@

As cryptocurrencies are based on free libre software, anyone can create a new currency.
If a currency is used by a single individual it is not very useful.
The potential of networked cryptocurrencies materializes when they are used by communities.
The communities can create or agree in a currency for their internal exchanges.
As a result, the currency is impregnated by the values and ethical standards of the community.
In a sense, the value of cryptocurrencies is backed only by the community of users.

@@

The value of the currency increases as the number of users and transactions increases.
Therefore, the members of a community have an incentive to devote some currency resources to attract new members to the community.
It is in their self-interest to make donations that increase the value of the currency.
In this regards, cryptocurrencies are a transition step towards a gift economy.

The gift (or reputation) economy is the kind of economy we find on the Internet, in which the participants offer software, articles, music, pictures and all kinds of digital goods for free.
This can be achieved using permissive licenses such as Creative Commons.
In the information age money is just information and wealth is measured in knowledge instead of material goods.
The progressive evolution towards a reputation economy is a natural step.

@@

Another example of the successful application of the community-centric model is the free libre open source movement.
Thousands of programmers affiliate to and contribute to this movement working as a self-managed community.
The software produced by this community often outperforms what have been developed using hierarchical models.
By making it possible for anyone to contribute, FLOSS projects have an advantage with respect to those that choose more restrictive models.

@@

\subsection{Commons for Europe}
Commons for Europe is a project focused on stimulating the production of commons resources in the areas of applications and networks.
The branch of Commons for Europe related to application is called Code for Europe and invites cities to incorporate programmers to their teams.
These programmers bring community and collaborative dynamics to the host institution and write software that later can be shared with other cities.

@@

\subsection{Confine}
The Confine project \cite{braem2013crc} emphasizes the links between research and community networks.
The goal is to make it easier for researchers to run experiments in community networks.
It provides the tools for automatize and monitor the experiments, as well as the hardware distributed across the community network.

@@

Networked cryptocurrencies are another example of community-centric innovation.
These currencies are developed by the communities as a result of open debate and discussion.
There are many similitudes with community networks.
The users of networked cryptocurrencies can contribute one node to the network to help in maintaining security.
The members of the community can also participate in many other ways offering products and services and using the cryptocurrency in their transactions.Networked cryptocurrencies

>>>>

Probably the best example of user-centric innovation are community networks.
Community networks are deployed by the same users that use them.
Consequently, they are carefully tailored to satisfy the users' needs.
These needs are not necessarily only about connectivity.
For example, the users might be interested in learning about the network.
Other aspects that can be of interest to the members of the community are the freedom, openness and neutrality of the network.
Some participants might be interested in researching  new services, applications and protocols.

@@

Another aspect that can draw the attention to community networks is privacy.
As commercial networks are involved in mass surveillance programs, the citizens see in community networks a way to opt out from privacy-invading campaigns.
Finally, community networks are also a means of getting around censorship.
Totalitarian regimes build censorship into the networks to control which information is distributed among the population.
Information filtering and Internet blackouts have been used as ways for hindering social movements and placate population unrest.

@@

A core principle of community networks is precisely the sharing of the network as a commons good.
This is particularly relevant in the case of wireless networks that use the ISM (industrial, scientific, medical) bands.
These bands contain a small fraction of the spectrum that is available without requiring a licence. 
In practice, these bands are a commons good that should be carefully managed by the community to use it effectively without creating unnecessary interference.
Good practices include the use of directive antennas and the limitation of the radiated power.

@@

There is a large number of community networks and they are very heterogeneous \cite{avonts2013qbe}.
They use different technologies such as wireless point-to-point links, wireless multi-point to multi-point links, ADSL or fiber.
Even in the case in which two community networks use the same layer-2 technology, it is still likely that they use different hardware, firmware or routing protocols.
The sizes of community networks also vary greatly and can range from few tens of nodes to tens of thousands.
The amount of documentation, supporting applications and organizational structure is also very different in different community networks.

Despite the disparity of models for community networks, there is an aspect that is common to all of them.
All these networks are community-centric.
They are planned, financed, deployed and maintained by the same users that benefit from them.

@@

Peer-to-peer networks are extremely resilient.
There is no single point of failure or central entity that can be attacked to take down the network.
As there are thousands of nodes spread across the network, the currency can survive as long as the underlying data network survives.
The risks of blockades or political interference are greatly mitigated.
Many Internet organizations such as Wikileaks and Wordpress accept donations and payments in cryptocurrency.
In some cases the reason to bypass government or paying networks blockades.
In others, the motivation is to operate in countries with lacking payment networks.

It is in areas lacking effective financial and banking systems where cryptocurrencies have a potential for greater impact.
Just like some areas leapfrogged from no telecommunication services to wireless handheld Internet enabled devices, it is possible that these same areas, that currently use mobile calls minutes as a currency, transition directly to the latest currency technology which is cryptocurrency.
Many countries suffer from chronic hyperinflation problems that could be solved by moving to a cryptocurrency in which the inflation rate (if any) can be chosen by the community and embedded in the protocol.

@@

As cryptocurrencies are based on free libre software, anyone can create a new currency.
If a currency is used by a single individual it is not very useful.
The potential of networked cryptocurrencies materializes when they are used by communities.
The communities can create (or agree in) a currency for their internal exchanges.
As a result, the currency is impregnated by the values and ethical standards of the community.
In a sense, the value of cryptocurrencies is backed only by the community of users.

@@

The value of the currency increases as the number of users and transactions increases.
Therefore, the members of a community have an incentive to devote some currency resources to attract new members to the community.
It is in their self-interest to make donations that increase the value of the currency.
In these regards, cryptocurrencies are a transition step towards a gift economy.

The gift (or reputation) economy is the kind of economy we find on the Internet in which the participants offer software, articles, music, pictures and all kinds of digital goods for free.
This can be achieved using permissive licenses such as Creative Commons.
In the information age, money is just information and wealth is measured in knowledge instead of material goods.
The progressive evolution towards a reputation economy is a natural step.

@@

Another example of the successful application of the community-centric model is the free libre open source software (FLOSS) movement.
Thousands of programmers affiliate to and contribute to this movement working as a self-managed community.
The software produced by this community often outperforms what have been developed using hierarchical models.
By making it possible for anyone to contribute, FLOSS projects have an advantage with respect to those that choose more restrictive models.

@@

\subsection{Commons for Europe}
Commons for Europe is a project focused on stimulating the production of commons resources in the areas of applications and networks.
The branch of Commons for Europe related to applications is called Code for Europe and invites cities to incorporate programmers to their teams.
These programmers bring community and collaborative dynamics to the host institution and write software that later can be shared with other cities.

@@

\subsection{Confine}
The Confine project \cite{braem2013crc} emphasizes the links between research and community networks.
The goal is to make it easier for researchers to run experiments in community networks.
It provides the tools for automatizing and monitoring the experiments, as well as the hardware that runs the experiments.
These hardware comprises Confine nodes which are distributed across the community network.

@@

Networked cryptocurrencies are another example of community-centric innovation.
These currencies are developed by the communities as a result of open debate and discussion.
There are many similitudes with community networks.
The users of networked cryptocurrencies can contribute one node to the network to help in maintaining security.
The members of the community can also participate in many other ways offering products and services and using the cryptocurrency in their transactions.Networked cryptocurrencies are also resilient to government interference or blockade.
They preserve the freedom of the user to spend the money in whatever she believes appropriate.

Community-centric innovation offers today tools that greatly empower the communities.
This opportunity can be captured to further evolve the existing tools and develop new ones that can facilitate the freedom and well-being of the population.
